Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition

### 2011-02-17 paper/main.tex
Added Allen's extensive changes.
old version moved to main_bak.tex
for temporary convenience, to be removed soon.

\usepackage{../authorkit/cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

@@

@@

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%\title{Accelerating $\ell_1$-Minimization Using Many-Core CPUs/GPUs \\ and Application to Face Recognition
%\title{Efficient Parallelization of Sparse Representation for Face Recognition}
%\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}
\title{Efficient Parallelization of $\ell_1$-Minimization for Face Recognition
\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}

@@

%Motivate the paper by an overview about the scope and importance of $\ell_1$-minimization.
$\ell_1$-minimization is one of the most powerful tools to be added to the engineer's
toolbox in years.  It has strong theoretical motivation for both for use in recovering sparse
solutions to systems of linear equations, and has been shown to perform well even in the
presence of a combination of sparse and dense error.  Since solutions of linear systems
arise in practically every branch of engineering and there are very few application which
do {\em not} benefit from a robust error function, the applications are far to numerous to
itemize here.  
For many applications of $\ell_1$-minimization solving a system $b=Ax$, $A$ often has a
special structure that can be leveraged to compute portions of $A$ on the fly
as they are needed, speeding up perfomance of the minimization dramatically.
Unfortunately, there are many import applications where this does not hold true.
This paper targets applications where $\bb$ and $A$ contain
image data, and thus must be loaded every time they are used, as is the 
case for many problems in computer vision.
The techniques presented in this paper target these applications, as well
as others where the dictionary is dense and unstructured. $A$ must be loaded every time
it is used, and algorithm performance becomes highly dependent on memory bandwidth, rather
than arithmetic logic.

% Narrow focus for applications to Recognition Application
The primary motivating application for this paper is face recognition for access control applications.
Recently, advances in automatic face recognition have been made by re-casting
it as a sparse representation problem.  This core of this technique consists of
re-casting recognition as a sparse representation problem. This is done by  
stacking the test image into a vector $\bb$, stacking the training images for
all of the users into the columns of a matrix $A$, and solving the following
minimization problem:
\begin{equation}
\min_{\x, \e} \| \x \|_1 + \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
\end{equation}
This optimization promotes a solution that is sparse both in $\x$ and $\e$. 
Sparsity in $\x$ arrises from the fact that only a subset of the training images (if any)
correspond to the user in the test image.  Sparsity in $\e$ arrises from the knowledge
that small occlusions in the test image will only corrupt a small subset of the pixels \cite{Wright2009-PAMI}.
The large coefficients in $\x$ will concentrate on the correct user, and can be used as the basis of a classifier.

For this idea to work, all of the images must be aligned. An iterative alignment
algorithm for this purpose was described as part of the \cite{Wagner2009-CVPR}, which was
the first paper to expand \cite{Wright2009-PAMI} into a functioning face recognition system.
This iterative alignment routine is based on repeatedly solving the following optimization problem:
\begin{equation}
\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A_k \x + \e.
\end{equation}
Where $A_k$ contains the training images for user $k$ as well as several vectors consisting of the
Jacobian of the test image $\bb$ with respect to the transformation parameters, and $\x$ contains
both the current representation coefficients, as well as an update to the transformation parameters.

\begin{figure}
\centering
%\includegraphics[width=3.5in]{figures/pipeline.pdf}
\input{figures/pipeline.tex}
\caption{The Full Recognition Pipeline}
\label{fig:pipeline}
\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=3.5in]{figures/alignment_l1.pdf}
%\caption{Alignment Stage L1 Solver}
%\label{fig:alignment_l1}
%\end{figure}

\subsection{Literature review}
%[Review the past parallel algorithms for $\ell_1$-minimization. Hopefully there aren't many]
%[Review the available different algorithms solving $\ell_1$-minimization and $\ell_0$-minimization, please refer to Allen's SIAM paper.
%[Zero in and justify why we choose ALM for our implementation.]

Standard solvers for $\ell_1$-minimization include orthogonal matching pursuit, basis pursuit and LASSO which cast the $\ell_1$-minimization as either a linear program or quadratic program.

These methods can be accelerated by more contemporary methods such as gradient projection, homotopy, iterative shrinkage-thresholding, proximal gradient, and augmented lagrangian methods.  According to analysis of these five algorithms applied to face recognition, Yang et al have determined that Homotopy and the Augmented Lagrangian Multiplier methods (ALM) achieve the best recognition rates while maintaining a low computational cost.  A closer look at these two algorithms will show that while the Homotopy method will produce the exact solution, the algorithm structure and functions do not lend themselves parallelization well.  
On the other hand, while the ALM finds an approximate solution, its computation mainly consists of matrix-vector operations, which lends itself to parallelization.

Past work in parallelizing $\ell_1$-minimization has been limited.  In 2009, Murphy at UCB implemented the L1-Spirit algorithm on Nvidia GPUs for MRI reconstruction and in 2009 Borghi at UCLA developed a proximal point algorithm using Moreau-Yosida regularization, which is special form of the $\ell_1$-minimization based on proximal gradient.

\subsection{Contributions of the paper}

\section{Augmented Lagrangian Multiplier Algorithm}

In this section, we describe the general version ALM algorithm and then describe how it can be adapted to solve our problem.

The augmented lagrange multiplier (ALM) method utilizes a popular class of convex techniques, the lagrange multiplier, solving the problem:
\begin{equation}
F(x) = f(x) + \lambda g(x).
\end{equation}
which can be formulated as an augmented Lagrange function given by
\begin{equation}
L_u(x,y) = \|x\|_1 + <\y, b - A\x - \e> + \frac{\mu}2 \| b-A\x-\e \|_2^2
\end{equation}

where $\mu > 0$ is a constant that penalizes infeasibility and $y$ is a vector of lagrange multipliers.

For an optimal $y*$ which satisfies the second-order sufficiency conditions, for a sufficiently large $mu$, it can be shown that
\begin{equation}
x* = \arg min L_u(x,y*)
\end{equation}

However, because $y*$ is generally unknown, ALM methods simultaneously estimate the optimal solution and Lagrange multipliers iteratively in the following procedure.

We also need to define the following soft-thresholding operator for a
scalar $x$ and a scalar $\alpha \geq 0$:
\begin{equation}
\textup{shrink}(x,\alpha) = \textup{sign}(x)\cdot \max \{|x| - \alpha, 0\},
\end{equation}
In practice, this operator will be applied elementwise to a vector $\x$ with a single $\lambda$,
and will be notated as $\textup{shrink}(\x,\alpha)$.

We summarize the entire ALM
algorithm as Algorithm~\ref{alg:alm}, where $\gamma$ denotes the largest eigenvalue of the matrix $A^TA$. For the choice of parameter $\mu$, we take the same strategy as
in \cite{YangJ2009-pp} and set $\mu_0 = 2m / \|\bb\|_1$. We set $\rho=1.5$.
\begin{algorithm}[h]
\caption{\bf (Augmented Lagrange Multiplier Method FIX)}
\begin{algorithmic}[1]
\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\blamda_1 = \mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $t_{1} \leftarrow 1, z_{1} \leftarrow \x_{k}, \u_{1} \leftarrow \x_{k}$;
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\e_{l+1} \leftarrow \textup{shrink}\left(\bb - A\x_l + \frac{\blamda_k}{\mu_k}, \frac{1}{\mu_k}\right)$;
\STATE $\x_{l+1} \leftarrow (A^\dagger)^T \left(\bb - \e_{l+1} + \frac{\blamda_k}{\mu_k} \right) $;
\ENDWHILE
\STATE $\blamda_{k+1} \leftarrow \blamda_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
\STATE $\mu_{k+1} \leftarrow \rho\mu_k$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{algorithmic}
\label{alg:alm}
\end{algorithm}

%In order to solve the alignment stage before recognition, we consider the $l_1$ problem: 
%\begin{equation} 
%\min_{\x, \e} \|\e\|_1 \quad \subj \quad \bb = A \x + \e.
%\end{equation}
%which is solved in a similar fashion.

% A note on alignment l1...
% With a reshuffling of the order of operations and a change of the termination
% condition, the inner loop takes the form of two matrix-vector multiplications
% and a sequence of vector-vector operations.

In order to adapt the general algorithm to face alignment and recognition, we formulated the 
$\ell_1$-minimization problem as described in \cite{}.

\subsection{Alignment Stage L1 Computational Complexity} Any discussion of the speed of an
algorithm should discuss the comutational complexity of the algorithm.  For the
ALM algorithms considered in this paper, this is rather simple.  The
Matrix-vector computions are $O(mn)$, and the vector-vector operations are
$O(m)$, in terms of both computation cost and memory-bandwidth.  Since A is
very tall, the computation of $A^\dagger$, the pseudoinverse of $A$, is best
computed via $A^\dagger = (A^TA)^{-1} A^T$.  The complexity of computing $G =
A^T A$ is $O(nnm)$.  $G$ is sufficiently small that it can be inverted via
Gauss-Jordan elimination, which has a complexity of $O(n^3)$.  The computation
of $A$ finishes with a matrix multiplication, with cost $O(n^2 m)$.
Fortunately, $A^\dagger$ only needs to be computed once per instance of the
$\ell_1$ minimization problem.

Therefore, for problems with large $n$, we expect the computation of the
pseudoinverse to dominate.  However, for the problem sizes needed in face
recognition problems, we find that the computation time is dominated by the
repeated computation of matrix-vector and matrix-matrix multiplications in the
inner loop.

\section{Hardware Parallelism} 
In this section we give a brief overview of the hardware architectures used
in this paper.

We feel it is most meaningful to compare hardware architectures on a per-board basis,
and in a typical hardware configuration.  GPU's are most commonly shipped with a single
GPU chip per PCI-express card.  
%For multicore CPU architectures, a typical workstation
%will consist of

For CPU implementations, the benchmarks makes use of all of
the cores in as many cpu's are present.  For GPU implementations, the benchmark
makes the best use of the entire GPU chip (most GPU boards have a single GPU
chip).  Unless otherwise specified, all implementations utilize single
precision floating point datatypes.

\subsection{CPU Hardware Parallelism} 
The baseline architecture for our experiments is a typical workstation,
with two quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each
processor has its own memory interface, and is directly connected to half of
the RAM installed in the machine.  The amount of RAM installed exceeds the
amount used by the algorithms, and is not an important performance
consideration.

Each core has a 32 KB L1 data cache and a 256 KB L2 cache. With Hyperthreading
enabled, each core can support two threads simultaneously.  For floating point
math, each core also has a vector unit capable of performing the same
arithmetic operation on four single precision floating point values
simultaneously.  \footnote{The core also has scalar floating point units, but
some compilers, notably gcc, generate code that uses the vector units
exclusively for floating point math.}

There are thus two important levels of parallelism that need to be exploited to
use a modern cpu efficiently: core-level parallelism and vector-level
parallelism.  For the implementations discussed here, we will primarily use the
OpenMP programming model to express core-level parallelism.  Depending on the
section of the code, we will achieve vector-level parallelism implicitly
through either the auto-vectorization features of Intel's icc compiler, or
through the use of a highly optimized BLAS library (Intel's MKL).

Each processor has 8MB of L3 cache that is shared by all four cores.  Overall,
this means the algorithm has approximately 16MB of on-chip cache available.
For the low resolution images used in sparse representation based face
recognition, this is enough cache to hold over 800 image-sized variables,
meaning that for small data sets it is possible to run completely using on-chip
cache.

@@

The ALM algorithm requires many matrix-vector operations, this gives manycore processors such as GPUs a computational advantage for large problem sizes.  GPUs are comprised of several streaming multiprocessors (SMP) (16 for the Nvidia Fermi) which each contain 32 floating-point units.  Each SMP contains its own L1-cache and shared memory and all SMPs share a common L2-cache.  The GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups of 32 threads called warps.  These warps are then grouped into higher levels called thread blocks and grids. The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance and a computational advantage over CPUs.  Unlike normal CPUs, the Nvidia Fermi has 768 kb of L1 cache/shared memory and 768kb of L2 cache.  This increase in performance doesn't come for free.  Because the GPU has it's own memory system, any data the GPU uses must be transferred from RAM to the GPU over PCI-Express, which has limited bandwidth.  However, the cost of these transfers are amortized by doing large amounts of computation at once, requiring code to be run completely on the GPU.

The GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.  Each thread block is a virtualized multiprocessor, which each thread as a virtualized scalar processor.  All of these differences allow GPUs to perform intensive calculations such as matrix-vector operations at much faster rates than CPUs.

% This section is Drew's
\section{Iterative Alignment Stage Implementation}

\subsection{Iterative Alignment on the CPU}
%\subsection{Tuning the GEMV operation}
%When matrix-vector operations are called sequentially on the same data,
%the operation can be blocked so that each thread operates on the same block of data for each operation.
%For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
%For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
%For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!

%\subsection{Tuning the Soft-Thresholding Operator}
%...Make sure there are no branches
%...Make sure there are no function calls
%...Make sure the loop gets automatically vectorized
%...Block the computation into work for different cpu cores
\subsection{Iterative Alignment on the GPU}
% Describingn the sequence of GPU implementations Drew has tried...

\subsection{Benchmark on Real Face Data}

%\begin{figure}
%\centering
%\includegraphics[width=3.5in]{figures/runtime_graph.pdf}
%\caption{GPU Alignment Stage Time vs. Size of Database}
%\label{fig:alignment_runtime}
%\end{figure}

\section{Recognition Stage Implementation}
% This section is Victor's

>>>>

\usepackage{iccv}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\def\arr{{\mathcal{A}}}
\newcommand{\Null}{{\mbox{Null}}}
\renewcommand{\mathbf}{\boldsymbol}
\renewcommand{\Re}{{\mathbb{R}}}
\def\tt{{\boldsymbol{t}}}
\def\ee{{\boldsymbol{e}}}
\def\aa{{\boldsymbol{a}}}
\def\bb{{\boldsymbol{b}}}
\def\dd{{\boldsymbol{d}}}
\def\ss{{\boldsymbol{s}}}
\def\cc{{\boldsymbol{c}}}
\def\pp{{\boldsymbol{p}}}
\def\chat{{\boldsymbol{\hat{\c}}}}
\def\nn{{\boldsymbol{n}}}
\def\rr{{\boldsymbol{n}}}
\def\m{{\boldsymbol{m}}}
\def\X{{\boldsymbol{X}}}
\def\L{{\boldsymbol{L}}}
\def\l{{\boldsymbol{l}}}
\def\Y{{\boldsymbol{Y}}}
\def\V{{\boldsymbol{V}}}
\def\uu{{\boldsymbol{u}}}
\def\vv{{\boldsymbol{v}}}
\def\yy{{\boldsymbol{y}}}
\def\xx{{\boldsymbol{x}}}
\def\ww{{\boldsymbol{w}}}
\def\zz{{\boldsymbol{z}}}
\def\uu{{\boldsymbol{u}}}
\def\F{{\boldsymbol{F}}}
\def\veronese{{\nu}}
\def\Qveronese{{\mu}}
\def\H{{\mathcal{H}}}
\def\Q{{\mathcal{Q}}}
\def\E{{\mathcal{E}}}
\def\L{{\mathcal{L}}}
\def\M{{\mathcal{M}}}
\def\A{{\mathcal{A}}}
\def\B{{\mathcal{B}}}
\def\F{{\mathcal{F}}}
\def\field{\mathbb{F}}
\def\QED{~\rule[-1pt]{5pt}{5pt}\par\medskip}

@@

\def\iccvPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ificcvfinal\pagestyle{empty}\fi
\begin{document}

%\title{Accelerating $\ell_1$-Minimization Using Many-Core CPUs/GPUs \\ and Application to Face Recognition
%\title{Efficient Parallelization of Sparse Representation for Face Recognition}
%\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}}
\title{Many-Core Parallelization of $\ell_1$-Minimization for Face Recognition
%\thanks{Corresponding author: . This work was partially supported by ARO MURI W911NF-06-1-0076.}
}

@@

$\ell_1$-minimization ($\ell_1$-min) has received much attention in recent years due to important applications
in compressive sensing \cite{BrucksteinA2007} and sparse representation \cite{WrightJ2010-PIEEE}.
The problem refers to finding the minimum $\ell$-norm solution to an underdetermined linear system $\bb=A\xx$:
\begin{equation}
\min \|\xx\|_1\quad \mbox{ subj. to }\quad \bb = A\xx.
\label{eq:l1min}
\end{equation}
It has now been well understood that, under certain conditions \cite{CandesE2005-IT_1,DonohoD2004}, the minimum $\ell_1$-norm solution is
also the \emph{sparsest} solution to the system \eqref{eq:l1min}.

Among its many applications, $\ell_1$-min has been recently used to reformulate image-based face recognition as a sparse representation problem \cite{WrightJ2009-PAMI}.
If we stack the training images of $K$ subject classes $(A_1\in\Re^{d\times n_1}, \cdots, A_K\in\Re^{d\times n_K})$ into the columns of a matrix $A = [A_1, \cdots, A_K]\in\Re^{d\times n}$, given a new query image in the same vector form $\bb\in\Re^d$, \emph{sparsity-based classification} (SBC) solves the following minimization problem:
\begin{equation}
\min_{\xx, \ee} \| \xx \|_1 + \|\ee\|_1 \quad \subj \quad \bb = A \xx + \ee.
\label{eq:l1min_denoise}
\end{equation}
If the sparsest solutions for $\x$ and $\e$ are recovered, $\ee$ provides a means to compensate corrupted pixels that main occlude some part of the query image, and the dominant nonzero coefficients in $\xx$ reveals the membership of $\bb$ based on the training image labels associated with $A$.

In this paper, we study parallelization of $\ell$-min on many-core CPU and GPU architectures. Firstly, although $\ell_1$-min \eqref{eq:l1min} is a convex program, conventional algorithms such as interior-point methods \cite{ChenS2001-SIAM,TibshiraniR1996} are known to suffer from poor scalability for large-scale applications such as face recognition. Recently, a number of accelerated algorithms have been proposed, which explicitly take advantage of the special structure of $\ell_1$-min problems \cite{LorisI2009,YangA2010-ICIP}. We investigate parallelization of a state-of-the-art $\ell_1$-min solution based on the classical framework of \emph{augmented Lagrange multiplier} (ALM) \cite{BertsekasD2003,YangA2010-ICIP}.

Secondly, while solving \eqref{eq:l1min_denoise} can achieve exceptional recognition accuracy on public databases, its success relies on the assumption that the query image is well aligned with the training images. In \cite{WagnerA2009-CVPR}, the framework of SBC was extended to iteratively align a query image to one subject class. The face alignment algorithm in turn solves the following minimization problem:
\begin{equation}
\hat{\tau}_i = \arg\min_{\xx, \ee, \tau_i} \|\ee\|_1\quad \mbox{subj. to}\quad \bb\circ\tau_i = A_i\xx + \ee,
\label{eq:l1min_alignment}
\end{equation}
where $\tau_i\in T$ is in a finite-dimensional group $T$ of transformations (e.g., affine or homography) acting on the image domain, which aligns $\bb$ with the training images from the $i$-th class $A_i$. The algorithm essentially extends Lucas-Kanade iterative alignment \cite{LucasB1981} to using the $\ell_1$ norm as a robust error function.
Although the original solution \cite{WagnerA2009-CVPR} was implemented based on standard interior-point methods, its speed has been improved by ALM in a more recent paper \cite{WagnerA2011-PAMI}. In this paper, in addition to solving the generic $\ell_1$-min objectives \eqref{eq:l1min} or \eqref{eq:l1min_denoise}, we also discuss how to properly customize the implementation to accelerate the face alignment step \eqref{eq:l1min_alignment} on many-core CPUs and GPUs.

Finally, we present extensive benchmark to compare the performance of the generic ALM $\ell_1$-min algorithm and the corresponding face recognition routines on massively parallel CPUs and GPUs. 
To this end, the face recognition system has been fully implemented on a state-of-the-art workstation, as shown in Figure \ref{fig:pipeline}. Since any algorithm parallelization is highly dependent on the specific computer hardware, the architecture of our test system will be explained in detail in Section \ref{sec:parallelism}.
We show that a highly accurate and robust face recognition system can be constructed to recognize face images from a moderate database of several hundred subjects. Note that with the possible continual increase of the number of subject classes, it is possible to modularize and dispatch different stages of the face recognition routine to a network of workstations to form an even larger scale networked system. Nevertheless, this scenario is not the focus of this paper. The source code of our algorithms, which is based on BLAS and Nvidia CUDA libraries, will be made available on our website to aid peer evaluation.
\begin{figure}
\centering
{\tiny \input{figures/pipeline_simplified.tex}}
\caption{The face recognition pipeline implemented on massively parallel hardware.}
\label{fig:pipeline}
\end{figure}

\subsection{Literature Review}
In this section, we briefly review the literature of $\ell_1$-min and existing works on its parallelization. Traditionally, $\ell_1$-min (or the so-called basis pursuit (BP)) has been formulated as a linear program \cite{ChenS2001-SIAM}. Several variations of the solution are also well known in optimization, including a noisy approximation via quadratic programming called the LASSO \cite{TibshiraniR1996} and truncated Newton interior-point method (TNIPM) \cite{KimS2007}.

One of the drawbacks of most interior-point methods for $\ell_1$-min (including BP, LASSO, and TNIPM) is that they require the solution sequence to follow an interior path, which is often computationally expensive. To mitigate these issues, an approach called \emph{Homotopy} has been studied to accelerate the speed of $\ell_1$-min. Homotopy methods for recovering sparse signals were first studied in the context of LASSO \cite{OsborneM2000}, which inspired a solution to the \emph{forward stagewise linear regression} problem called LARS \cite{EfronB2004}, and eventually led to the corresponding Homotopy algorithms for BP in \cite{MalioutovD2005,DonohoD2006}.

Although Homotopy can be shown to exactly estimate BP when the solution is sufficiently sparse \cite{DonohoD2006}, the algorithm still involves expensive matrix operations such as matrix-matrix multiplication and solving least-squares problems with varying $A$ matrices. Hence, it is not efficient to implement Homotopy on parallel hardware. In Section \ref{sec:ALM}, we contend that ALM is a better choice among Homotopy and other existing $\ell_1$-min solvers to parallelize on many-core CPUs and GPUs. The algorithm belongs to a category of approximate $\ell_1$-min solutions called \emph{iterative shrinkage-thresholding} (IST) method \cite{WrightS2008,BeckA2009}. The (sequential) implementation of IST algorithms mainly involves elementary operations such as vector algebra and matrix-vector multiplication. Therefore, when the dimension of the problem becomes high, IST-type algorithms are particularly suitable for systems in parallel-computing environments. In \cite{YangA2010-ICIP}, the authors further showed that, compared to other IST-type algorithms, ALM is able to achieve good estimation accuracy comparable to BP and Homotopy, while at the same time significantly improve the speed compared to these exact $\ell_1$-min solutions. Due to these facts, in this paper, we choose ALM as the core algorithm to implement the complete parallel face recognition pipeline.

In terms of the past works in parallel $\ell_1$-min, the literature has been limited, to the best of our knowledge. In \cite{BorghiA2010}, Borghi et al. developed a special proximal gradient $\ell_1$-min algorithm based on Moreau-Yosida regularization. In \cite{MurphyM2010}, Murphy et al. presented parallel implementation of the l1-SPIRIT MRI reconstruction algorithm on NVidia programmable GPUs.

%\subsection{Contributions of the Paper}

\section{Augmented Lagrange Multiplier Method}
\label{sec:ALM}

For the completeness of the paper, in this section, we briefly describe the ALM algorithm for $\ell_1$-min \eqref{eq:l1min} \cite{YangA2010-ICIP} and analyze its complexity. Parallelization of ALM for face alignment and recognition will be discussed in Section \ref{sec:alignment} and Section \ref{sec:recognition}, respectively.

Lagrange multiplier methods have been frequently used in convex programming to eliminate equality constraints via adding a significant penalty term to the cost function for infeasible points. ALM methods differ from other penalty-based approaches by simultaneously estimating the optimal solution and Lagrange multipliers in an iterative fashion.
For $\ell_1$-min, the augmented Lagrange function is defined as:
\begin{equation}
L_\mu(\xx,\yy) = \|\xx\|_1 + \yy^T(\bb - A\xx - \ee) + \frac{\mu}{2} \| \bb-A\xx-\ee \|_2^2,
\end{equation}
where $\mu > 0$ is a constant that penalizes infeasibility and $\yy$ is a vector of lagrange multipliers.

In Lagrange Multiplier Theory, if there exists a Lagrangian $y^*$ that satisfies the second-order sufficiency conditions for optimality \cite{BertsekasD2003}, then for a sufficiently large $mu$, the optimal $\ell_1$-min solution also minimizes
\begin{equation}
\xx^* = \arg \min L_\mu(\xx,\yy^*).
\label{eq:optimal-ALM}
\end{equation}

In practice, the optimal values for the triplet $(\xx^*, \yy^*, \uu)$ are all unknown. Furthermore, it has been observed that if solving \eqref{eq:optimal-ALM} with a large initial value of $\mu$ tends to lead to slower convergence speed \cite{WrightS2008,YangA2010-ICIP}. In \cite{BertsekasD2003,YangJ2009}, an alternating procedure has been proposed to iteratively update $\xx$ and $\yy$:
\begin{equation}
\left \{
\begin{array}{lll}
\xx_{k+1} & = & \arg\min_{\xx} \, L_{\mu_k} (\xx,\yy_k)\\
\yy_{k+1} & = & \yy_k + \mu_k (\bb - A\xx_{k+1}) \\
\end{array}
\right . ,
\label{eq:alm}
\end{equation}
where $\left\{\mu_{k}\right\}$ is a monotonically increasing positive sequence. The iteration terminates when the estimates $(\xx, \yy)$ converges.

Note that in the iterative procedure outlined in \eqref{eq:alm}, the second step only involves vector algebra and matrix-vector multiplication. Therefore, the procedure is computationally efficient only if it is easier to minimize the augmented Lagrange function compared to solving the original problem \eqref{eq:l1min} directly. In fact, the first step that minimizes $L_{\mu_k} (\xx,\yy_k)$ assuming $\yy_k$ and $\mu_k$ are fixed can be solved element-wise iteratively by a soft-thresholding algorithm \cite{WrightS2008,BeckA2009}, whose time complexity is bounded by $O(n^2)$ and can be easily parallelized. Algorithm \ref{alg:alm} summarizes the ALM $\ell_1$-min algorithm.
 \begin{algorithm}[h]
\caption{Augmented Lagrange Multiplier (ALM)}
{\bf INPUT:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$, $\tau\leftarrow$ largest eigenvalue of $A^TA$, and a constant $\rho>1$.
\begin{algorithmic}[1]
\WHILE{not converged ($k = 1,2,\ldots$)} \STATE $t_1 \leftarrow 1$, $\zz_1
\leftarrow \xx_k$, $\uu_1 \leftarrow \xx_k$ \WHILE{not
converged ($l = 1,2,\ldots$)} \STATE $\uu_{l+1}  \leftarrow
\mathrm{soft}\left(\zz_l - \frac{1}{\tau}A^T\left(A\zz_l - \bb
- \frac{1}{\mu_k}\yy_k\right), \frac{1}{\mu_k\tau}\right)$
\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
\sqrt{1+4t_l^2}\right)$ \STATE $\zz_{l+1} \leftarrow \uu_{l+1}
+ \frac{t_l - 1}{t_{l+1}}(\uu_{l+1} - \uu_l)$ \ENDWHILE \STATE$
\xx_{k+1} \leftarrow \uu_{l+1}$ \STATE $\yy_{k+1} \leftarrow
\yy_k + \mu_k (\bb - A\xx_{k+1})$ \STATE $\mu_{k+1} \leftarrow
\rho\cdot\mu_k$ \ENDWHILE 
\end{algorithmic}

{\bf OUTPUT:} $\xx^* \leftarrow \xx_k$.
\label{alg:alm}
\end{algorithm}

\section{Hardware Parallelism}
\label{sec:parallelism}
In this section we discuss the hardware architectures used in this paper. Aside from the $\ell_1$-min algorithm, our choice
of the architectures plays a significant role in many practical decisions we make in Section \ref{sec:alignment} and Section \ref{sec:recognition}.
For CPU implementation, the system would make use of all the cores in as many CPUs are present on one motherboard.  For GPU implementation, the system would
make the best use of a single many-core Nvidia GPU chip  on a PCI card (GPUs are commonly shipped with a single GPU chip per PCI-express card).  Unless otherwise specified, all implementations utilize single
precision floating point datatypes.

\subsection{CPU Hardware Parallelism} 
The baseline architecture for our experiments is a Linux workstation with two quad-core Intel Nehalem E5530 processors clocked at 2.4 GHz.  Each
processor has its own memory interface, and is directly connected to half of
the RAM installed in the machine.  The amount of RAM installed exceeds the
amount used by the algorithms, and is not an important performance
consideration.

Each core has a 32\,KB L1 data cache and a 256\,KB L2 cache. With Hyperthreading
enabled, each core can support two threads simultaneously.  For floating point
instructions, each core also has a vector unit capable of performing the same
arithmetic operation on four single precision floating point values
simultaneously.  \footnote{The core also has scalar floating point units, but
some compilers, notably gcc, generate code that uses the vector units
exclusively for floating point instructions.}

There are thus two important levels of parallelism that need to be exploited to efficiently
use a modern CPU: core-level parallelism and vector-level
parallelism.  For the implementations discussed here, we will primarily use the
OpenMP programming model to express core-level parallelism {\bf AY: What is OpenMP? citation needed}.  Depending on the
section of the code{\bf AY: What is a section of the code?}, we will achieve vector-level parallelism implicitly
through either the auto-vectorization features of Intel's icc compiler, or
through the use of a highly optimized BLAS library (Intel's MKL). {\bf AY: citations needed, what is icc, what is the relation between BLAS and MKL? Are they equivalent? not clear in the context.}

Each processor further has 8MB of L3 cache that is shared by all four cores.  Overall,
this means the algorithm has approximately 16MB of on-chip cache available.
For the low resolution images used in sparse representation based face
recognition, this is enough cache to hold over 800 image-sized variables, {\bf AY: how low resolution? which sparse representation problem? ``image-sized variables'' means simply ``image vectors''?}
meaning that for small data sets it is possible to run completely using on-chip
cache.

@@

The ALM algorithm requires many matrix-vector operations, this gives many-core processors such as GPUs a computational advantage for large problem sizes.  GPUs are comprised of several streaming multiprocessors (SMP) (16 for the Nvidia Fermi) which each contains 32 floating-point units.  Each SMP contains its own L1-cache and shared memory and all SMPs share a common L2-cache.  The GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups of 32 threads called \emph{warps}.  These warps are then grouped into higher levels called \emph{thread blocks} and \emph{grids}. The GPU contains an inverted memory hierarchy so that the L1 cache stays close to the cores, providing the increase in performance and a computational advantage over CPUs.  Unlike normal CPUs, the Nvidia Fermi has 768 kb of L1 cache/shared memory and 768kb of L2 cache. However, this increase in performance does not come for free. As the GPU has its own memory system, any data the GPU uses must be transferred from RAM to the GPU over PCI-Express, which has limited bandwidth. However, the cost of these transfers are amortized by doing large amounts of computation at once, requiring code to be run completely on the GPU.

In summary, a programmable GPU can be analogous to a multi-core CPU processor, except with more sophisticated vector processing hardware.  Each thread block is a virtualized multiprocessor, which each thread as a virtualized scalar processor.  All of these differences allow GPUs to perform intensive calculations such as matrix-vector operations at much faster rates than CPUs.

% This section is Drew's
\section{Parallelization of the Face Alignment Stage}
\label{sec:alignment}
{\bf AY: Add an algorithm here describe the alignment block in the pipeline. Explain the algorithm and necessary modification from generic l-1}

\subsection{CPU Implementation}

{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
%\subsection{Tuning the GEMV operation}
%When matrix-vector operations are called sequentially on the same data,
%the operation can be blocked so that each thread operates on the same block of data for each operation.
%For Single threaded MKL sgemv, Code took          0.0335823 msec per instance to run!
%For Automatic multi-threaded MKL sgemv, Code took 0.008179 msec per instance to run!
%For Manually multi-threaded MKL sgemv, Code took  0.0061276 msec per instance to run!

%\subsection{Tuning the Soft-Thresholding Operator}
%...Make sure there are no branches
%...Make sure there are no function calls
%...Make sure the loop gets automatically vectorized
%...Block the computation into work for different cpu cores
\subsection{GPU Implementation}
{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}

\section{Parallelization of the Face Recognition Stage}
\label{sec:recognition}
{\bf AY: Add an algorithm here describe the recognition block in the pipeline. Explain the algorithm and necessary modification from generic l-1}

\subsection{CPU Implementation}
{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}

\subsection{GPU Implementation}
{\bf AY: A diagram should be used to illustrate how the jobs are distributed on a CPU architecture.}
The GPU ALM algorithm is written in CUDA and utilizes the CUBLAS libraries.
CUBLAS are the BLAS libraries consisting of all single precision BLAS calls
which are provided by Nvidia and are optimized for GPU performance. In
addition, we coded kernel (GPU) methods for vector-vector operations to replace
operations that would have taken multiple BLAS calls. Because data transfer from the RAM to GPU suffers bandwidth limitations, we designed our implementation to transfer the A matrix and measurement vector at once.  The algorithm is similar
to the CPU algorithm except that we had to parallelize the code at an
additional level. The CPU portion of the code serves as the conductor, making
calls to GPU functions and directing the flow of the code while the GPU keeps all of the data and runs the computation.

\section{Experiment}
{\bf AY: Describe the benchmark protocol here. Both for simulation and real face data.}

\subsection{Simulation}

\subsection{Face Recognition Benchmark}
